---
title: "Economic Connectedness"
author: Alfred Cheung
format: html
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
library(tidyverse)
library(primer.data)
```


The histogram illustrates the distribution of U.S. county populations in 2018, based on data from Opportunity Insights. Displayed on a logarithmic scale, the x-axis reveals a highly skewed distribution, where most counties have populations between 10,000 and 100,000, while a smaller number of counties have populations exceeding one million. The use of a log scale helps visualize this imbalance, making it easier to observe patterns among the smaller counties that make up the majority. This skewed distribution highlights the importance of accounting for population differences when analyzing county-level data, particularly in research related to economic connectedness, social capital, and policy design.

```{r}
connectedness |> 
  drop_na() |> 
  ggplot(aes(x = population)) +
    geom_histogram(bins = 100) +
    scale_x_log10(labels = scales::comma_format()) +
    labs(title = "US County Populations in 2018",
         subtitle = "Populations are highly skewed",
         x = "Population",
         y = "Number of Counties",
         caption = "Opportunity Insights")
```



This scatter plot visualizes the relationship between U.S. county population size and economic connectedness, using data from the Opportunity Insights connectedness dataset. Each point represents a county, with the x-axis showing population (on a logarithmic scale) and the y-axis measuring economic connectedness—a metric that captures how closely individuals in low-income households are socially connected to those in higher-income groups. The plot includes a linear trend line, which appears nearly flat, indicating that there is little to no correlation between population size and economic connectedness across counties. This suggests that larger counties do not necessarily offer stronger economic connections for low-income individuals, and that smaller counties can be just as, or even more, economically connected. The use of a log scale allows for better visibility across the wide range of population sizes, helping reveal this subtle but important pattern.

```{r}
ggplot(connectedness, aes(x = population, y = connectedness)) +
  geom_point(color = "black", alpha = 0.7, size = 1.5) +
  geom_smooth(method = "lm", color = "blue", se = FALSE, linewidth = 1) +
  scale_x_log10(labels = scales::comma_format()) +
  labs(
    title = "US County Populations versus Economic Connectedness",
    subtitle = "Population size does not matter to EC",
    x = "Population",
    y = "Economic Connectedness",
    caption = "Opportunity Insights"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.caption = element_text(hjust = 1),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(margin = margin(b = 10))
  )
```
